---
title: "COVID Deaths"
author: "Sean L"
date: "27/08/2021"
output: html_document
---

## COVID deaths are very much in the public spotlight right now.

Let's focus in on proportions of those deaths and see where most of the deaths are occuring

First let's bring in our libraries

```{r}
library(tidyverse) # import
library(dplyr) # import
library(ggplot2)
```

We need the tidyverse so we can mess around with our data and graph it and such. next, let's read in our data from a cvs file

```{r}
data <- read.csv("C:/Users/seanl/Desktop/Portfolio projects/deaths_covid_data.csv", header=TRUE)

```

Now that our data is in, I notice that in our file there are a ton of blank cells in the deaths columns. That will become a problem when we try to sum everything up and maximize and such. Let us change all of those to zeros.

```{r}
data$total_deaths[is.na(data$total_deaths)] <- 0
head(data)
```

Now the total deaths column is all zero's where the blank cells used to be. I could go through and do this for all of them, but for my analysis I only need the total deaths column to work out.

Let's make a new dataframe that gets the maximum deaths for each country.

```{r}
total_deaths_by_country <- data %>% group_by(location) %>% summarise(total_deaths = max(total_deaths, na.rm = TRUE))


View(total_deaths_by_country)
```

And there we are.Let's add a new column and find the proportion of world deaths each country had.

```{r}
total_deaths_by_country$death_ratio <- ((total_deaths_by_country$total_deaths/4441418)*100) 
head(total_deaths_by_country)
```

Also notice that this data Frame has continents and the world total in as well. Let's remove those temporarily...

```{r}
total_deaths_by_country <- total_deaths_by_country[-c(2, 12, 13, 69, 154, 194, 230),]
View(total_deaths_by_country)
```

And now we'll make a pie chart showing all the countries.

```{r}
ggplot(total_deaths_by_country, aes(x = "", y = location, fill = total_deaths)) +
  geom_bar(stat = "Identity") +
  coord_polar(theta = "y")
```

Okay, even if I wanted to clean that up it's really ugly. There are just too many countries. But we can use this idea to make a pie chart for each continent and see their contributions to total covid deaths.

```{r}
total_deaths_by_continent <- total_deaths_by_country[c(2, 12, 13, 69, 154, 194),]
View(total_deaths_by_continent)
```

```{r}
ggplot(total_deaths_by_continent, aes(x = 1, y = total_deaths, fill = location))+
  geom_col(position = 'stack') +
  geom_text(aes(label = paste(location, ': ', total_deaths)), 
            position = position_stack(vjust = .5)) +
  coord_polar(theta = "y") +
  theme_void()
```

